$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-03-29 paper/e6ssm-tuning-paper.tex
more fixes due to corrected U(1)_I results.

However we should also acknowledge that there are other limits which
play a similar role.  Chargino limits directly constrain the $\mu$
parameter (or effective $\mu$ parameter in these $U(1)$ extensions).
However the chargino limits from the LHC depend on whether there are
light sleptons or sneutrinos and the mass difference between the
lightest chargino and lightest neutralino.  Current limits placed by
CMS and ATLAS extend up to $m_{\tilde{\chi}^\pm_1} \approx 700-740$ GeV if
there are light sleptons\cite{Khachatryan:2014qwa,Aad:2014nua} with
much weaker bounds if there are no light sleptons or
sneutrinos\footnote{Useful summary plots of these limit may be found
  on the public pages of ATLAS,
  \url{https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/CombinedSummaryPlots/SUSY/ATLAS_SUSY_EWSummary/ATLAS_SUSY_EWSummary.png}
  and CMS
  \url{http://cms.web.cern.ch/sites/cms.web.cern.ch/files/styles/large/public/field/image/Image_03_exclusion_Combined.png?itok=8FMBpu_1}.}.

@@

Interestingly the coefficient $d$ is very small (and negative) in the
case of the $U(1)_I$.  This allows a dramatic reduction in the fine
tuning from the $U(1)_I$ symmetry.  This is a result of the $H_u$
charge associated with $U(1)_I$ vanishing, which means that the
D-terms to the lightest Higgs which is predominantly $H_u$ at large
$\tan \beta$ are suppressed, making it difficult to raise the Higgs
mass in the same way as happens in the other models and explaining why
heavier Higgs values in this model can't be obtained.  Therefore the fine
tuning behaviour in this model is closer to that of the MSSM, though
the LHC can still constrain the most natural scenarios from just
placing tougher limits on the $Z^\prime$ mass.

@@

In contrast we find that in $U(1)$ extensions of the MSSM there is an
additional way to constrain the naturalness of the models, which is
through the $Z^\prime$ mass limit.  We find when we impose a low
cutoff of $20$ TeV for setting the soft masses, the lowest tuning in
the E$_6$SSM is compatible with a $Z^\prime$ mass of $2.5$ TeV was
$\Delta \approx 121$, while if the LHC run II can place a limit of
$4.5$ TeV on $M_Z^\prime$ then the tuning would be approximately $394$.
By comparison the current situation in the MSSM only requires a tuning
of around $38$. This should be interpreted as saying that in
the most conservative limits one can place on naturalness on these
models, the tuning in the E$_6$SSM is worse.  However if there are no
charginos below $700$ GeV then the situation in the two models would
be the same.

@@

We also looked at the tuning in different $U(1)$ extensions for
fixed $\tan \beta = 10$.  We found that in all cases the fine tuning
was much worse for the larger $Z^\prime$ mass, further emphasising
the importance of this in $U(1)$ extensions.  Interestingly
the actual level in tuning varied quite a bit between these models
with the $U(1)_I$ model showing the least tuning due to the vanishing
charge of the $H_u$ state. This model is quite interesting in the
sense that it provides a solution to the $\mu$-problem while avoiding
the large tuning (with current limits) from the $Z^\prime$
mass. However one should remember we are looking at conservative limits
on naturalness here and there is no solution to the usual tuning coming from
the large stops needed to get a $125$ GeV Higgs in this model.  In any
case future LHC limits on the $Z^\prime$ mass would be especially
interesting for this case as they would significantly raise the
minimum level of tuning which can be obtained.

>>>>

However we should also acknowledge that there are other limits which
play a similar role.  Chargino limits directly constrain the $\mu$
parameter (or effective $\mu$ parameter in these $U(1)$ extensions).
However the chargino limits from the LHC depend on whether there are
light sleptons or sneutrinos and the mass difference between the
lightest chargino and lightest neutralino.  Current limits placed by
CMS and ATLAS extend up to $m_{\tilde{\chi}^\pm_1} \approx 700-740$ GeV if
there are light sleptons\cite{Khachatryan:2014qwa, Aad:2014nua} with
much weaker bounds if there are no light sleptons or
sneutrinos\footnote{Useful summary plots of these limit may be found
  on the public pages of ATLAS,
  \url{https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/CombinedSummaryPlots/SUSY/ATLAS_SUSY_EWSummary/ATLAS_SUSY_EWSummary.png}
  and CMS
  \url{http://cms.web.cern.ch/sites/cms.web.cern.ch/files/styles/large/public/field/image/Image_03_exclusion_Combined.png?itok=8FMBpu_1}.}.

@@

Interestingly the coefficient $d$ is very small (and negative) in the
case of the $U(1)_I$.  This allows a dramatic reduction in the fine
tuning from the $U(1)_I$ symmetry.  This is a result of the $H_u$
charge associated with $U(1)_I$ vanishing, which means that the
D-terms to the lightest Higgs which is predominantly $H_u$ at large
$\tan \beta$ are suppressed, making it difficult to raise the Higgs
mass in the same way as happens in the other models and explaining why
heavier Higgs values in this model can't be obtained.  Therefore the
fine tuning behaviour in this model is closer to that of the MSSM, and
in this case raising the $Z^\prime$ mass limit to $4.5$ TeV will have
little impact on naturalness.  From naively estimating the tuning
using the $d$ coeffioceint one can estimate that $Z^\prime$ limits
need to be around $15$ TeV before they will raise the tuning in this
model.

@@

In contrast we find that in $U(1)$ extensions of the MSSM there is an
additional way to constrain the naturalness of the models, which is
through the $Z^\prime$ mass limit.  We find when we impose a low
cutoff of $20$ TeV for setting the soft masses, the lowest tuning in
the E$_6$SSM is compatible with a $Z^\prime$ mass of $2.5$ TeV was
$\Delta \approx 121$, while if the LHC run II can place a limit of
$4.5$ TeV on $M_Z^\prime$ then the tuning would be approximately $394$.
By comparison the current situation in the MSSM only requires a tuning
of around $38$. This should be interpreted as saying that in
the most conservative limits one can place on naturalness in these
models, the tuning in the E$_6$SSM is worse.  However if there are no
charginos below $700$ GeV then the situation in the two models would
be the same.

@@

We also looked at the tuning in different $U(1)$ extensions for fixed
$\tan \beta = 10$.  We found that in every case except for the
$U(1)_I$ the fine tuning was much worse for the larger $Z^\prime$
mass, further emphasising the importance of this in $U(1)$ extensions.
The $U(1)_I$ model showed the least tuning due to
the vanishing charge of the $H_u$ state. This model is quite
interesting in the sense that it provides a solution to the
$\mu$-problem while avoiding the large tuning (with current limits)
from the $Z^\prime$ mass. However one should remember we are looking
at conservative limits on naturalness here and there is no solution to
the usual tuning coming from the large stops needed to get a $125$ GeV
Higgs in this model.
